---
title: "COVID-19 Canada"
runtime: shiny
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(lubridate)
library(incidence)
knitr::opts_chunk$set(warning=FALSE)
# Sys.unsetenv("http_proxy")
# Sys.unsetenv("https_proxy")
```


## Covid 19 Epidemionlogy with R

Rf.: 

- the following analysis is based on this [post](https://rviews.rstudio.com/2020/03/05/covid-19-epidemiology-with-r/)
- You can find [Modelling the effects of public health interventions on COVID-19 transmission using R - part 1](https://timchurches.github.io/blog/posts/2020-03-10-modelling-the-effects-of-public-health-interventions-on-covid-19-transmission-part-1/) in details
- also the [R Epidemics Consortium](https://www.repidemicsconsortium.org/projects/) will provide you the tool

### Getting data

We are now getting data from JHU data gathered since beginning of the COVID-19. The address below contains the data live. We are going to subset it to Canada.

```{r}
jhu_url <- paste("https://raw.githubusercontent.com/CSSEGISandData/", 
  "COVID-19/master/csse_covid_19_data/", "csse_covid_19_time_series/", 
  "time_series_covid19_confirmed_global.csv", sep = "")
jhu_url
```

```{r}
ca_confirmed_long_jhu <- read_csv(jhu_url) %>% 
  rename(province = "Province/State", 
  country_region = "Country/Region") %>% 
  pivot_longer(-c(province, country_region, Lat, Long), 
               names_to = "Date", 
               values_to = "cumulative_cases") %>% 
  # adjust JHU dates back one day to reflect US time, more or less
  mutate(Date = mdy(Date) - days(1)) %>% 
  filter(country_region == "Canada") %>% 
  arrange(province, Date) %>% 
  group_by(province) %>% 
  mutate(
    incident_cases = c(0, diff(cumulative_cases))
    ) %>% 
  ungroup() %>% 
  select(-country_region, -Lat, -Long) %>% 
  filter(str_detect(province, "Diamond Princess", negate = TRUE))

ca_confirmed_long_jhu
```

Plot the cumulative incidence increase per province:

```{r}
ca_confirmed_long_jhu %>%
  ggplot(aes(x=Date, y=cumulative_cases, color=province)) +
  geom_line() 
```

### Calculate at country level

So calculate the cumulative and incident cases at group level:


```{r}
ca_confirmed <- ca_confirmed_long_jhu %>%
  group_by(Date) %>%
  summarise(
    cumulative_cases = sum(cumulative_cases),
    incident_cases = sum(incident_cases)
  )

ca_confirmed %>%
  ggplot(aes(x=Date, y=cumulative_cases)) +
  geom_line() 

require(scales)
ca_confirmed %>%
  ggplot(aes(x=Date, y=cumulative_cases)) +
  geom_line()  +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks()  
```

### A quick fit exponential models to incidence data

```{r}
ca_confirmed %>%
  filter(Date >= as_date("2020-02-27")) %>%
  select(Date, incident_cases) %>%
  uncount(incident_cases) -> ca_incidence

ca_incidence_object <- incidence(ca_incidence$Date, last_date = max(ca_incidence$Date))
ca_incidence_object
ca_incidence_fit <- incidence::fit(ca_incidence_object, quiet = FALSE)
ca_incidence_fit$info$pred
```



```{r}
plot(ca_incidence_object) %>% 
  add_incidence_fit(ca_incidence_fit)
```

### Estimate R0 


```{r}
library(earlyR)
library(epitrix)
mu <- 7.5  # days
sigma <- 3.4  # days
param <- epitrix::gamma_mucv2shapescale(mu, sigma/mu)

w <- distcrete::distcrete("gamma", interval = 1, shape = param$shape, scale = param$scale, 
    w = 0)
w
R <- get_R(ca_incidence_object, si = w, si_mean=mu, si_sd=sigma)
get_R(ca_incidence_object, si_mean=mu, si_sd=sigma)
R

growth_R0 <- lm2R0_sample(ca_incidence_fit$model, w) 

growth_R0 %>% summary()
growth_R0 %>% hist(col = "grey", border = "white", main = "Distribution of R0")
```

```{r}
plot(R)
```

```{r}
plot(R, "lambdas")
```

### Project the future


```{r}
library(projections)

pred_fwd_days <- 7

pred_growth <- project(ca_incidence_object, R = median(growth_R0), si=w, n_sim=10000, n_days = pred_fwd_days)

plot(ca_incidence_object) %>%
  add_projections(pred_growth) %>% 
  add_incidence_fit(ca_incidence_fit) +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  ggtitle('Incidence per day with projection')



pred_cum <- cumulate(pred_growth) # cumulative predictions

plot(cumulate(ca_incidence_object)) %>%
  add_projections(pred_cum+ca_incidence_object$n) +
  ggtitle('Cumulative incidence projection')+
  theme(axis.text.x=element_text(angle=90,hjust=1))


plot(pred_cum+ca_incidence_object$n) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 20))
```

### Estimate Time Varying Reproduction Numbers from Epidemic Curves

The mean and std is from WHO estimation of serial interval.

```{r fig.height=10}
library(EpiEstim)

days <- seq_along(ca_incidence_object$dates)

mu <- 7.5  # days
sigma <- 3.4  # days

Rt <- estimate_R(ca_incidence_object, 
                 method = "parametric_si",
                  config = make_config(list(
                  mean_si = mu, std_si = sigma)))

# custom results plotting function to avoid the ugly
# TableGrob messages returned by the plotting function in the
# EpiEstim package
plot_Ri <- function(estimate_R_obj) {
    p_I <- plot(estimate_R_obj, "incid", add_imported_cases = TRUE)  # plots the incidence
    p_SI <- plot(estimate_R_obj, "SI")  # plots the serial interval distribution
    p_Ri <- plot(estimate_R_obj, "R")
    return(gridExtra::grid.arrange(p_I, p_SI, p_Ri, ncol = 1))
}

plot_Ri(Rt)

```

We now put serial interval to be uncertain and simulate the results.

```{r fig.height=10}
ca_res_uncertain_si <- estimate_R(ca_incidence_object, method = "uncertain_si", 
    config = make_config(list(mean_si = 7.5, std_mean_si = 2, 
        min_mean_si = 1, max_mean_si = 8.4, std_si = 3.4, std_std_si = 1, 
        min_std_si = 0.5, max_std_si = 4, n1 = 100, n2 = 100)))

plot_Ri(ca_res_uncertain_si)


```

We estimate the R0 at daily level.

```{r fig.height=10}
t_start <- seq(2, length(ca_incidence_object$dates)-1)
t_end <- t_start + 1


empirical_si_config_daily <-  make_config(list(mean_si = 7.5, std_mean_si = 2, 
        min_mean_si = 1, max_mean_si = 8.4, std_si = 3.4, std_std_si = 1, 
        min_std_si = 0.5, max_std_si = 4, n1 = 100, n2 = 100, t_start = t_start, t_end = t_end))

ca_res_uncertain_si_daily <- estimate_R(ca_incidence_object, 
    method = "uncertain_si", config = empirical_si_config_daily)


plot_Ri(ca_res_uncertain_si_daily)
```

